---
title: "R Assignment 1- Neighbourhood Crime Rate"
Inttroduction: The data describes crimerate in Toronto region in 2020. It shows the number of diffrent crimes in comparion to neighbourhood population.
output: pdf_document
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[CO,CE]{Group4 - Assignment R}
---
## Installing packages and library
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggpubr)
library(tidyverse)
library(janitor) # janitor helps us clean datasets
library(here) # here helps us to know where files are
library(readxl)
```

## Load data - CSV
## This is a Neighbourhood Crime Rate dataset and downloaded Open data

```{r crimerate}
crimedata <- readr::read_csv(file ="C:/Users/Dell/Downloads/Neighbourhood_Crime_Rates (1).csv")


```

## 2.Print the structure of your dataset.



```{r}
str(crimedata)
```
##3.List the variables in your dataset.



```{r}
names(crimedata)
```
##4.Print the top 15 rows of your dataset.
```{r}
head(crimedata, n=15)
```



##5. Write a user defined function using any of the variables from the data set.
```{r}
Crime_2019_2020 <-function(a,b,c){a*b+c}
Crime_2019_2020(crimedata$Assault_2020,crimedata$AutoTheft_2020,crimedata$BreakAndEnter_2020)
```
##6.Use data manipulation techniques and filter rows based on any logical criteria that exist in your dataset.



```{r}
over_25000 <- filter(crimedata,crimedata$F2020_Population_Projection >25000)
print(over_25000)
```
## 7.Identify the dependent & independent variables and use reshaping techniques and create a new data frame by joining those variables from your dataset.



```{r}
rate1 <- as.data.frame (crimedata %>% select(6))
rate2 <-as.data.frame (crimedata%>% select(7))
r1_r2 = data.frame(rate1,rate2)
print(r1_r2)
```

#8. Remove missing values in your dataset. • 

```{r}
crimedata1 = na.omit(`crimedata`) 
```
 


#9. Identify and remove duplicated data in your dataset. 

#Identify - 
```{r}
crimedata[duplicated('crimedata')] 

```


#Remove -  
```{r}
crimedata1[!duplicated(crimedata1$Neighbourhood),]
```


#10.Reorder multiple rows in descending order
#  ex) due to double dataset, using first dataset for descending order.
```{r}
crimedata %>% arrange(desc(F2020_Population_Projection))
```



#11.Rename some of the column names in your dataset
```{r}
library(plyr)
library(dplyr)
renamecrimedata <- crimedata
colnames(renamecrimedata)
names(renamecrimedata)[names(renamecrimedata) == "F2020_Population_Projection"] <-"Pop_Proj_2020"
names(renamecrimedata)[names(renamecrimedata) == "Neighbourhood"] <- "Region"
view(renamecrimedata)
print(renamecrimedata)
```


#12.Add new variables in your data frame by using a mathematical function 
#   (for e.g. –multiply an existing column by 2 and add it as a new variable to your data frame)
#Top 10 Neighbourhood with highest crimerate
```{r}
Crimerate1 <- crimedata %>% mutate(Total_Crimerate= round(((Assault_2020+AutoTheft_2020
+BreakAndEnter_2020+Robbery_2020+Shootings_2020)/F2020_Population_Projection
)*100000,digits=2))%>% arrange(desc(Total_Crimerate))
head(Crimerate1, 10)
print(head (Crimerate1,10))
view(head (Crimerate1,10))

```


#13. Plot a scatter plot for any 2 variables in your dataset.
```{r}

library(ggplot2)

ggplot(data = crimedata, aes(x = F2020_Population_Projection, y=  AutoTheft_Rate2020 ))+ geom_point(size =1.5, color = "red", shape = 10)
```


#14. Plot a bar plot for any 2 variables in your dataset. 
#geom_bar() or geom_col()
```{r}
library(ggplot2)
ggplot(data = crimedata, aes(x= BreakAndEnter_2020))+geom_bar(fill = "aquamarine4")
```


#15. Find the correlation between any 2 variables by applying least square linear regression model.
```{r}
ggplot(data = crimedata, aes(x = Assault_Rate2020, y= BreakAndEnter_Rate2020))+geom_smooth()
ggplot(data = crimedata, aes(x = Assault_Rate2020, y= BreakAndEnter_Rate2020))+ geom_boxplot()
```





Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#16.using any of the numerical variables from the dataset and perform the following statistical functions mean, median, mode and range

#Mean
```{r}
mean(crimedata$Assault_2020)
```

#Median
```{r}
median(crimedata$Assault_2020)
```


 
#Mode
```{r}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}


v <- c(crimedata$Assault_2020)


result <- getmode(v)
print(result)

```

#Range
```{r}
range(crimedata$Assault_2020)
```

#17. Create a training set using random number generator engine
```{r}
head(crimedata)
h <-runif(nrow(crimedata))
crime<-crimedata[order(h), ]
str(crime)
train <- crime[1:50, ] 
crime.data1 = na.omit(crimedata)
set.seed(1234)
trainingcrimedata = as.data.frame(crime.data1 %>% sample_frac(0.75,replace = FALSE))

```

#18.Print the summary statistics of your dataset
```{r}
summary(crimedata)
```

#19.Conclusion: We can Conclude that The top crime prone neighbourhood in Toronto in 2020 was Moss Part followed by Kensington-Chinatown and Church-Yonge Corridor. Breaking and entering was much less compared to assaults.






